---
title: "Getting started with the robis package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{getting-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

First load the package:

```{r setup}
library(robis)
```

## Occurrences

The `occurrence()` function provides access to raw occurrence data. For example, to fetch all occurrences by scientific name:

```{r message = FALSE}
occurrence("Abra aequalis")
```

Alternatively, occurrences can be fetched by AphiaID:

```{r message = FALSE}
occurrence(taxonid = 293683)
```

Other parameters include `geometry`, which accepts polygons in WKT format:

```{r message = FALSE}
occurrence("Abra alba", geometry = "POLYGON ((2.59689 51.16772, 2.62436 51.14059, 2.76066 51.19225, 2.73216 51.20946, 2.59689 51.16772))")
```

A convenience function `map_leaflet()` is provided to visualize occurrences on an interactive map:

```{r eval = FALSE}
map_leaflet(occurrence("Abra sibogai"))
```

## Checklists

The `checklist()` function returns all taxa observed for a given set of filters.

```{r message = FALSE}
checklist("Semelidae")
```

Just like the `occurrence()` function, `checklist()` accepts WKT geometries:

```{r message = FALSE}
checklist(geometry = "POLYGON ((2.59689 51.16772, 2.62436 51.14059, 2.76066 51.19225, 2.73216 51.20946, 2.59689 51.16772))")
```

## Measurements and facts

The package also provides access to `MeasurementOrFact` records associated with occurrences. When calling `occurrence()`, `MeasurementOrFact` records can be included by setting `mof = true`.

```{r message = FALSE}
occ <- occurrence("Abra tenuis", mof = TRUE)
```

`MeasurementOrFact` records are nested in the occurrence, but the `measurements()` function allows you to extract them to a flat data frame. Use the `fields` parameter to indicate which occurrence fields need to be preserved in the measurements table.

```{r message = FALSE}
mof <- measurements(occ, fields = c("scientificName", "decimalLongitude", "decimalLatitude"))
mof
```

Note that the `MeasurementOrFact` fields can be used as parameters to the `occurrence()` function. For example, to only get occurrences with associated biomass measurements:

```{r message = FALSE}
library(dplyr)

occurrence("Abra tenuis", mof = TRUE, measurementtype = "biomass") %>%
  measurements()
```
